---
title: "2015-3-2-MattGhazarian-HW2"
author: "Matthew Ghazarian"
date: "February 25, 2015"
output: html_document
---
##Setup
``` {r}
# require libraries
require(reshape2)
require(ggplot2)
require(plyr)
require(scales)

# read csv file
raw.data <- read.csv("Microbe-scope-bugs.csv")
```

##Cleaning
```{r}
#select 5 columns that I need
mydata <- subset(raw.data, select = c(X, case.fatality.rate, primary.mode.of.transmission, average.basic.reproductive.rate))

#rename colnames using plyr rename function
df <- rename(mydata, c("X"="name", "case.fatality.rate"="fat","primary.mode.of.transmission"="trans", "average.basic.reproductive.rate"="R0" ))

#erase the row 1 notes and row 2 anthrax (not on the chart because missing values)
df <- df[3:41, 1:4]

#remove unneeded percentage signs
df$fat <-gsub( "%", "", paste(df$fat))

#set fatalities percentages as numbers so I can plot them
df$fat <- as.numeric(df$fat)

#set R0 to numbers -- this needs to be converted from a factor to a character, first, to avoid unwanted changes to the values, since factors are stored as integers
df$R0 <- as.numeric(as.character(df$R0))

#delete "Malaria (P. falciparum)" because our target project from Information Is Beautful also deleted it (likely because it's an outlier on the R0 scale)
df <- df[-c(16), ]

#convert trans column from factors to chracters
df$trans <- as.character(df$trans)

#combine "airborne droplet" and "airborne" categories to match McCandless' graph, which also combines them
df$trans <-gsub( "airborne droplet", "airborne", paste(df$trans))

#assign "airborne" to cholera, which seems to accidentally have been left blank, but is marked "airborne" in McCandless' map
df[6,3] <- "airborne"
```




##Plotting with ggplot 
Since the first plot is quite messy with overlapping labels, and since I could not figure out how to execute a discontinuous y-axis soon enough, I tried ways to clean up the appearance without the discontinuous scale. Still, jittering techniques, and other methods I attempted could not clean up the overlapping labels before the assignment was due, so I have resorted to creating a second plot for the low-level fatalities. 

In my defense, Hadley Wickham is also hesitant of discontinuous y-axes, as they distort the data. He talks about this topic [here](https://groups.google.com/forum/#!topic/ggplot2/jSrL_FnS8kc)
```{r}
#first plot to see what it looks look
p1 <- ggplot(df, aes(x=R0, y=fat)) + geom_point(position= position_jitter(w = 2, h= 2), aes(color=trans)) +
  geom_text(aes(label=name, color=trans, size=2.5, angle=30)) +
  guides(color = guide_legend("Primary transmission method")) + 
  guides(size=FALSE)

#change x and y axes labels
p1 <- update_labels(p1, list
        (x = "Contagiousness (average basic reproduction number, R0)", 
        y="Deadliness (case fatality rate)"))

#add main title 
p1 <- p1 + labs(title = "The Microbe-scope (copy)")

#print
p1

#create dataframe of low level values
df0 <- subset(df, fat <= 1)

#plot them separately
p2 <- ggplot(df0, aes(x=R0, y=fat)) +
  geom_point(position = position_jitter(w = 0.2, h = 0.0), aes(color=trans)) +
  geom_text(aes(label=name, color=trans, size=2.5, angle=30, vjust=2)) +
  guides(color = guide_legend("Primary transmission method")) +
  guides(size=FALSE)

#provide new title and change labels
p2 <- update_labels(p2, list
        (x = "Contagiousness (average basic reproduction number, R0)", 
        y="Deadliness (case fatality rate)"))
p2 <- p2 + labs(title = "The Microbe-scope (copy), low-fatality cases")

#print
p2                 
```

